---
title: "p8105_hw1_hx2306"
author: "HuijunXiao"
date: "9/21/2021"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggplot2)
```

## Problem 1   

* Step 1: Create the dataframe   

```{r}
# a random sample of size 10 from a standard Normal distribution
nordist_sam <-
  rnorm(10)

# a logical vector indicating whether elements of the sample are greater than 0
sam_greater0 <-
  nordist_sam >0 

# a character vector of length 10
chara_10 <-
  c("a","b","c","d","e","f","g","h","i","j")

# a factor vector of length 10, with 3 different factor “levels”
fact_10 <-
  c("good","bad","ok","good","bad","ok","good","bad","ok","good")%>%
  as.factor()
```

```{r}
# problem 1 data frame
df_problem1 <-
  tibble(nordist_sam,sam_greater0,chara_10,fact_10)

df_problem1
```

* Step 2: Take the mean of each variable in the data frame   

```{r}
# Mean of nordist_sam
mean(pull(df_problem1,nordist_sam))

# Mean of sam_greater0
mean(pull(df_problem1,sam_greater0))

# Mean of chara_10
mean(pull(df_problem1,chara_10))

# Mean of fact_10
mean(pull(df_problem1,fact_10))

```

__Only numeric and logical variables are able to take the means.For character and factor variables, taking the mean does not work and will return NA instead.__

* Step 3: Add `as.numeric` function to logical,character and factor variables   

```{r, results="hide"}
# add as.numeric to nordist_sam
as.numeric(pull(df_problem1,nordist_sam))

# add as.numeric to sam_greater0
as.numeric(pull(df_problem1,sam_greater0))

# add as.numeric to chara_10
as.numeric(pull(df_problem1,chara_10))

# add as.numeric to fact_10
as.numeric(pull(df_problem1,fact_10))
```

**`as.numeric` function can be applied to logical variables successfully.**  
**However, it cannot be applied to character and factor variables.**  
**This helps explain what happens when I try to take the mean as factor and character strings are not properly formatted numbers.**   
**Hence, they cannot be converted to the numeric class and means cannot be calculated as well.**   

## Problem 2   

```{r}
#Load Penguins dataset 
data("penguins", package = "palmerpenguins")
```

* Step 1: Short description of the `penguins` dataset   

```{r}
# the data in this dataset, including names / values of important variables
head(penguins)
```
**`penguins` dataset contains `r ncol(penguins)` variables, which are species, island, bill length, bill depth, flipper length, body mass, sex and year, and `r nrow(penguins)` rows of observations.**  

**`penguins` dataset included data recorded from `r min(pull(penguins,year), na.rm = T)` to `r max(pull(penguins,year), na.rm = T)`.**  

**For species of this dataset, there are `r length(unique(penguins$species))` categories, which are `r unique(penguins$species)`.** 

**For island of this dataset, there are `r length(unique(penguins$island))` categories, which are `r unique(penguins$island)`.**

**The bill length (mm) starts from `r min(pull(penguins,bill_length_mm), na.rm = T)` to `r max(pull(penguins,bill_length_mm), na.rm = T)` with mean equals to `r mean(pull(penguins,bill_length_mm), na.rm = T)`.**

**The bill depth (mm) starts from `r min(pull(penguins,bill_depth_mm), na.rm = T)` to `r max(pull(penguins,bill_depth_mm), na.rm = T)` with mean equals to `r mean(pull(penguins,bill_depth_mm), na.rm = T)`.** 

**The flipper length (mm) starts from `r min(pull(penguins,flipper_length_mm), na.rm = T)` to `r max(pull(penguins,flipper_length_mm), na.rm = T)` with mean equals to `r mean(pull(penguins,flipper_length_mm), na.rm = T)`.** 

**The body mass (g) starts from `r min(pull(penguins,body_mass_g), na.rm = T)` to `r max(pull(penguins,body_mass_g), na.rm = T)` with mean equals to `r mean(pull(penguins,body_mass_g), na.rm = T)`.** 
  
* Step 2: Make a scatterplot of flipper_length_mm (y) vs bill_length_mm (x)
  
```{r}
## make plot
ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species)) + 
  geom_point() +
  ggtitle("Penguins Flipper Length (mm) vs. Bill Length (mm)") +
  ylab("Flipper Length (mm)") +
  xlab("Bill Length (mm)")

## save plot
ggsave("p8105_hw1_hx2306_scatterplot.jpeg",plot = last_plot())
```

